---
title: "lab01"
output: html_document
date: "2022-11-09"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

1. Check the results of the snippets. Comment what is going on.

```{r Q1, echo=TRUE}
x1 <- 1/3 ; x2 <- 1/4
if ( x1 - x2 == 1/12) {
  print("Subtraction is correct")
} else {
  print("Subtraction is wrong")
}

# and

x1 <- 1 ; x2 <- 1/2
if ( x1 - x2 == 1/2) {
  print("Subtraction is correct")
} else {
  print("Subtraction is wrong")
}
```

- 1/3 and 1/12 are infinite decimals, while 1/2 is not.  
Computation of floating-point numbers makes the result different.

2. If there are any problems, suggest improvements.
- give a tolerance |a - b| < ε， while ε is close to 0
```{r}
x1 <- 1/3 ; x2 <- 1/4
if ( abs((1 / 12) - (x1 - x2)) < exp(-30)) {
  print("Subtraction is correct")
} else {
  print("Subtraction is wrong")
}
```

or directly using all.equal()
```{r}
x1 <- 1/3 ; x2 <- 1/4
if (all.equal(1 / 12, x1 - x2) ) {
  print("Subtraction is correct")
} else {
  print("Subtraction is wrong")
}
```


## Question 2

1. Write your own R function to calculate the derivative of $f(x) = x$ in this way with $ε = 10^{-15}$
```{r}
Q2_derivative <- function(x){
  
  e <- 10^-15
  fp <-  ((x + e) - x) / e
  
  return(fp)
}

```

2. Evaluate your derivative function at x = 1 and x = 100000.
```{r}
Q2_derivative(1)
Q2_derivative(100000)
```

3. What values did you obtain? What are the true values? Explain the reasons behind the
discovered differences.
- True value should be exactly 1. While what we got is 1.110223 at x = 1 and 0 at x = 100000.  
ε + x - x makes floating-point number rounded. And due to 100000 occupies too much mantissa, the result was rounded to 0.


## Question 3

```{r}
# 1
myvar <- function(x){
  n <- length(x)
  s2 <- (sum(x^2) - (sum(x))^2 / n) / (n - 1)
  
  return(s2)
}

# 2
x <- rnorm(10000, mean = 10^8, sd = 1)

myvar(x)

# 3
Y <- numeric(length = 10000)
for (i in 1:10000) {
  Y[i] <- myvar(x[1:i] - var(x[1:i]))
}

plot(1:10000, Y)
plot(1:10000, Y, xlim = c(1,50))
```



