---
title: "Lab06"
author: "Shipeng Liu,Dongwei Ni"
date: "`r Sys.Date()`"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggplot2)
library(dplyr)
```

# Assignment 1: Genetic Algorithm

## Task 1
```{r}
fx=function(x){
  ((x^2)/exp(x))-2*exp(-(9*sin(x))/(x^2+x+1))
}

```


## Task 2:Deﬁne the function crossover()
```{r}
crossover=function(x,y){
  return((x+y)/2)
}

```


## Task 3:Deﬁne the function mutate()
```{r}
mutate=function(x){
  return((x^2)%%30)
}
```


## Task 4:

```{r}
genetic=function(maxiter,mutprob){
  X_init=seq(0,30,5)
  X=X_init
  values=sapply(X,FUN=fx)
  for(i in 1:maxiter){
    parents=sample(X,2)
    victim=order(values)[1]
    kid=crossover(parents[1],parents[2])
    prob=runif(1,0,1)
    if(prob<mutprob){
      kid=mutate(kid)
    }
    X[victim]=kid
    values=sapply(X,FUN=fx)
  }
  
  
  #plot
  plot_dataframe=data.frame("x"=seq(0,30,0.001),
                            "fx"=sapply(seq(0,30,0.001),FUN=fx),
                            "class"="function_plot")
  plot_dataframe=rbind(plot_dataframe,data.frame("x"=X,
                                                 "fx"=sapply(X,FUN=fx),
                                                 "class"="observation"))
  p=ggplot(data=plot_dataframe)+geom_point(aes(x=x,y=fx,color=class))
  plot(p)
  
  return(X)
}

```

## Task 5: Run the code with diﬀerent combinations of maxiter= 10, 100 and mutprob= 0.1, 0.5, 0.9.

### maxiter=10,mutprob=0.1
```{r}
X_observe=genetic(10,0.1)
cat("The final observations is:",X_observe)
```

### maxiter=10,mutprob=0.5
```{r}
X_observe=genetic(10,0.5)
cat("The final observations is:",X_observe)
```

### maxiter=10,mutprob=0.9
```{r}
X_observe=genetic(10,0.9)
cat("The final observations is:",X_observe)
```

### maxiter=100,mutprob=0.1
```{r}
X_observe=genetic(100,0.1)
cat("The final observations is:",X_observe)
```

### maxiter=100,mutprob=0.5
```{r}
X_observe=genetic(100,0.5)
cat("The final observations is:",X_observe)
```

### maxiter=100,mutprob=0.9
```{r}
X_observe=genetic(100,0.9)
cat("The final observations is:",X_observe)
```


### Conclusion:

Among this case,we could find that when maxiter is larger,and mutprob is smaller,the optimizer will perform better.The best optimizer in these cases is the conbination where maxiter=100,mutprob=0.1.


# Assignment 2: EM algorithm

## Task 1:Make a time series plot
```{r}
physical=read.csv("physical1.csv")

ggplot(physical)+geom_line(aes(x=X,y=Y),color="blue")+
  geom_line(aes(x=X,y=Z),color="red")+ylab("Values")
```

It seems that two processes are related to each other,the variance of both Y and Z decrease as the X increase.


# Task 2:


# Task 3:

```{r}
loglikelihood=function(lamb){
  
  
}

EM<-function(Y,eps){
    lambd_init=100
    lambd_curr=lambd_init
    while (abs(lambd_prev-lambd_curr)<0.001){
	lambd_prev<-lambd_curr
	## E-step
	EY<-sum(Yobs)+(n-r)*muk
	EY2<-sum(Yobs^2)+(n-r)*(muk^2+sigma2k)

	## M-step
	muk<-EY/n
	sigma2k<-EY2/n-muk^2

	## Compute log-likelihood
	llvalcurr<-floglik(Yobs,muk,sigma2k,r)
	k<-k+1

	print(c(muk,sigma2k,llvalcurr))
    }
}

```
















































